Fat--Tailed Variational Inference with Anisotropic Tail Adaptive Flows

### 2022-08-23 design/equivalents_nips/neurips_2020.tex
Updates from Overleaf

\documentclass{article}

@@

% to avoid loading the natbib package, add option nonatbib:
\ifisarxiv
\usepackage{fullpage}
\usepackage{natbib}
\else
\usepackage[final]{neurips_2020}
\fi
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{hyperref}       % hyperlinks
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography
 \usepackage{subfigure}
\usepackage{amsmath,amsthm}
\usepackage{cancel}
\usepackage{cleveref}
\usepackage{mathtools}
\usepackage{enumitem}
\usepackage{todonotes}
\usepackage{xcolor}
\usepackage{xfrac}
\usepackage{caption}
\hypersetup{
	colorlinks,
	linkcolor={red!40!gray},
	citecolor={blue!40!gray},
	urlcolor={blue!70!gray}
}
\newcommand{\BLUE}{\color[rgb]{0,0,0.69}}
\input{../shortdefs}

@@

\begin{document}
\maketitle

\begin{abstract}
It is often desirable to reduce the dimensionality of a large dataset
by projecting it onto a low-dimensional subspace.  Matrix sketching
has emerged as a powerful technique for performing such dimensionality
reduction very efficiently.  Even though there is an extensive
literature on the worst-case performance of sketching, existing
guarantees are typically very different from what is observed in
practice.  We exploit recent developments in the spectral analysis of
random matrices to develop novel techniques that provide provably
accurate expressions for the expected value of random projection
matrices obtained via sketching.  These expressions can be used to
characterize the performance of dimensionality reduction in a variety
of common machine learning tasks, ranging from low-rank approximation
to iterative stochastic optimization.  Our results apply to several
popular sketching methods, including Gaussian and Rademacher sketches,
and they enable precise analysis of these methods in terms of spectral
properties of the data.  Empirical results show that the expressions
we derive reflect the practical performance of these sketching
methods, down to lower-order effects and even constant factors. 
\end{abstract}

@@

Matrices that have known spectral decay, most commonly with either
exponential or polynomial rate, arise in many machine
learning problems \cite{randomized-newton}. Such behavior can be naturally
occurring in data, or it can be induced by feature expansion using, say, the
RBF kernel (for exponential decay) \cite{Santa97Gaussianregression} or the
Mat\'ern kernel (for polynomial decay) \cite{RasmussenWilliams06}. Understanding these
two classes of decay plays an important role in distinguishing the
properties of light-tailed and heavy-tailed data distributions.
Note that in the kernel setting we may often represent our data via
the $m\times m$ kernel matrix $\K$, instead of the $m\times n$ data
matrix $\A$, and study the sketched Nystr\"om method
\cite{revisiting-nystrom} for low-rank approximation. To handle 
the kernel setting in our analysis, it suffices to replace the squared singular
values $\sigma_i^2$ of $\A$ with the eigenvalues of $\K$.
\begin{figure}[t]
\centering
\subfigure[%
Singular values are given by $\sigma_i^2=C\cdot \alpha^{i-1}$.]{%
    \includegraphics[width=0.48\textwidth]{explicit_exp}
  }
\hfill
\subfigure[%
Singular values are given by $\sigma_i^2=C\cdot i^{-\beta}$.
]{%
      \includegraphics[width=0.48\textwidth]{explicit_poly}
    }
\caption{
Theoretical predictions of low-rank approximation error of a Gaussian
sketch under known spectral decays, compared to the empirical results.
The constant $C$ is scaled so that $\|\A\|_F^2=1$ and we let
$n=m=1000$. For the theory, we plot the explicit formulas
\eqref{eq:explicit-exp} and \eqref{eq:explicit-poly} (dashed lines),
as well as the implicit expression from Corollary \ref{c:low-rank}
(thin solid lines) obtained by numerically solving
\eqref{eq:equation}. Observe that the explicit and implicit
predictions are nearly (but not exactly) identical.
}
%\vspace{-5mm}
\label{f:explicit}
\end{figure}

@@

\bibliographystyle{alpha}
\bibliography{../pap}

@@

\ifisarxiv\else\newpage\fi
\appendix

>>>>

% \documentclass{article}
\documentclass[thesis.tex]{subfiles}

@@

% to avoid loading the natbib package, add option nonatbib:
\ifisarxiv
% \usepackage{fullpage}
% \usepackage{natbib}
% \else
% \usepackage[final]{neurips_2020}
% \fi
% \usepackage[utf8]{inputenc} % allow utf-8 input
% \usepackage[T1]{fontenc}    % use 8-bit T1 fonts
% \usepackage{hyperref}       % hyperlinks
% \usepackage{url}            % simple URL typesetting
% \usepackage{booktabs}       % professional-quality tables
% \usepackage{amsfonts}       % blackboard math symbols
% \usepackage{nicefrac}       % compact symbols for 1/2, etc.
% \usepackage{microtype}      % microtypography
%  \usepackage{subfigure}
% \usepackage{amsmath,amsthm}
% \usepackage{cancel}
% \usepackage{cleveref}
% \usepackage{mathtools}
% \usepackage{enumitem}
% \usepackage{todonotes}
% \usepackage{xcolor}
% \usepackage{xfrac}
% \usepackage{caption}
% \hypersetup{
% 	colorlinks,
% 	linkcolor={red!40!gray},
% 	citecolor={blue!40!gray},
% 	urlcolor={blue!70!gray}
% }
% \newcommand{\BLUE}{\color[rgb]{0,0,0.69}}
% \input{../shortdefs}

@@

\begin{document}
% \maketitle

% \begin{abstract}
It is often desirable to reduce the dimensionality of a large dataset
by projecting it onto a low-dimensional subspace.  Matrix sketching
has emerged as a powerful technique for performing such dimensionality
reduction very efficiently.  Even though there is an extensive
literature on the worst-case performance of sketching, existing
guarantees are typically very different from what is observed in
practice.  We exploit recent developments in the spectral analysis of
random matrices to develop novel techniques that provide provably
accurate expressions for the expected value of random projection
matrices obtained via sketching.  These expressions can be used to
characterize the performance of dimensionality reduction in a variety
of common machine learning tasks, ranging from low-rank approximation
to iterative stochastic optimization.  Our results apply to several
popular sketching methods, including Gaussian and Rademacher sketches,
and they enable precise analysis of these methods in terms of spectral
properties of the data.  Empirical results show that the expressions
we derive reflect the practical performance of these sketching
methods, down to lower-order effects and even constant factors. 
% \end{abstract}

@@

Matrices that have known spectral decay, most commonly with either
exponential or polynomial rate, arise in many machine
learning problems \cite{randomized-newton}. Such behavior can be naturally
occurring in data, or it can be induced by feature expansion using, say, the
RBF kernel (for exponential decay) \cite{Santa97Gaussianregression} or the
Mat\'ern kernel (for polynomial decay) \cite{RasmussenWilliams06}. Understanding these
two classes of decay plays an important role in distinguishing the
properties of light-tailed and heavy-tailed data distributions.
Note that in the kernel setting we may often represent our data via
the $m\times m$ kernel matrix $\K$, instead of the $m\times n$ data
matrix $\A$, and study the sketched Nystr\"om method
\cite{revisiting-nystrom} for low-rank approximation. To handle 
the kernel setting in our analysis, it suffices to replace the squared singular
values $\sigma_i^2$ of $\A$ with the eigenvalues of $\K$.
\begin{figure}[t]
\centering
    \begin{subfigure}{0.49\textwidth}%
\subcaption{Singular values are given by $\sigma_i^2=C\cdot \alpha^{i-1}$.}
    \includegraphics[width=0.48\textwidth]{explicit_exp}
    \end{subfigure}
\hfill
    \begin{subfigure}{0.49\textwidth}%
\subcaption{Singular values are given by $\sigma_i^2=C\cdot i^{-\beta}$.}
      \includegraphics[width=0.48\textwidth]{explicit_poly}
    \end{subfigure}
\caption{
Theoretical predictions of low-rank approximation error of a Gaussian
sketch under known spectral decays, compared to the empirical results.
The constant $C$ is scaled so that $\|\A\|_F^2=1$ and we let
$n=m=1000$. For the theory, we plot the explicit formulas
\eqref{eq:explicit-exp} and \eqref{eq:explicit-poly} (dashed lines),
as well as the implicit expression from Corollary \ref{c:low-rank}
(thin solid lines) obtained by numerically solving
\eqref{eq:equation}. Observe that the explicit and implicit
predictions are nearly (but not exactly) identical.
}
%\vspace{-5mm}
\label{f:explicit}
\end{figure}

@@

% \bibliographystyle{alpha}
% \bibliography{../pap}

@@

% \ifisarxiv\else\newpage\fi
% \appendix

### 2022-08-23 ftvi/ftvi_icml/ftvi_icml.tex
Updates from Overleaf

\documentclass[nohyperref]{article}

% Recommended, but optional, packages for figures and better typesetting:
\usepackage{microtype}
\usepackage{graphicx}
%\usepackage{subfigure}
\usepackage{booktabs} % for professional tables

% hyperref makes hyperlinks in the resulting PDF.
% If your build breaks (sometimes temporarily if a hyperlink spans a page)
% please comment out the following usepackage line and replace
% \usepackage{icml2022} with \usepackage[nohyperref]{icml2022} above.
\usepackage{hyperref}

% Attempt to make hyperref and algorithmic work together better:
\newcommand{\theHalgorithm}{\arabic{algorithm}}

% Use the following line for the initial blind version submitted for review:
% \usepackage{icml2022}

% If accepted, instead use the following line for the camera-ready submission:
\usepackage[accepted]{icml2022}

% For theorems and such
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{mathtools}
    \usepackage{amsthm}

% if you use cleveref..
\usepackage[capitalize,noabbrev]{cleveref}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% THEOREMS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}[section]
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{assumption}[theorem]{Assumption}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Todonotes is useful during development; simply uncomment the next line
%    and comment out the line below the next line to turn off comments
%\usepackage[disable,textsize=tiny]{todonotes}
\usepackage[textsize=tiny]{todonotes}

% The \icmltitle you define below is probably too long as a header.
% Therefore, a short form for the running title is supplied here:
\icmltitlerunning{Fat--Tailed VI with Anisotropic Tail Adaptive Flows}

% custom stuff
\usepackage[bold]{hhtensor}
\usepackage{mathtools}
\usepackage{float}
\usepackage{braket}
\usepackage{dsfont}
\usepackage{graphbox}
\usepackage{pgf}
\usepackage{comment}
\usepackage{subcaption}
\usepackage{enumitem}
\input{../macros}

@@

\twocolumn[
\icmltitle{Fat--Tailed Variational Inference with Anisotropic Tail Adaptive Flows}

% It is OKAY to include author information, even for blind
% submissions: the style file will automatically remove it for you
% unless you've provided the [accepted] option to the icml2022
% package.

% List of affiliations: The first argument should be a (short)
% identifier you will use later to specify author affiliations
% Academic affiliations should list Department, University, City, Region, Country
% Industry affiliations should list Company, City, Region, Country

% You can specify symbols, otherwise they are numbered in order.
% Ideally, you should not use this facility. Affiliations will be numbered
% in order of appearance and this is the preferred way.
\icmlsetsymbol{equal}{*}

\begin{icmlauthorlist}
\icmlauthor{Feynman Liang}{yyy,fb}
\icmlauthor{Liam Hodgkinson}{yyy,zzz}
\icmlauthor{Michael W. Mahoney}{yyy,zzz}
% \icmlauthor{Firstname4 Lastname4}{sch}
% \icmlauthor{Firstname5 Lastname5}{yyy}
% \icmlauthor{Firstname6 Lastname6}{sch,yyy,comp}
% \icmlauthor{Firstname7 Lastname7}{comp}
% %\icmlauthor{}{sch}
% \icmlauthor{Firstname8 Lastname8}{sch}
% \icmlauthor{Firstname8 Lastname8}{yyy,comp}
% %\icmlauthor{}{sch}
% %\icmlauthor{}{sch}
\end{icmlauthorlist}

\icmlaffiliation{fb}{Meta, Menlo Park, CA}
\icmlaffiliation{yyy}{Department of Statistics, University of California, Berkeley, CA}
\icmlaffiliation{zzz}{International Computer Science Institute, Berkeley, CA}
% \icmlaffiliation{comp}{Company Name, Location, Country}
% \icmlaffiliation{sch}{School of ZZZ, Institute of WWW, Location, Country}

\icmlcorrespondingauthor{Feynman Liang}{feynman@berkeley.edu}
%\icmlcorrespondingauthor{Firstname2 Lastname2}{first2.last2@www.uk}

% You may provide any keywords that you
% find helpful for describing your paper; these are used to populate
% the "keywords" metadata in the PDF but will not be shown in the document
\icmlkeywords{Machine Learning, ICML}

\vskip 0.3in
]

% this must go after the closing bracket ] following \twocolumn[ ...

% This command actually creates the footnote in the first column
% listing the affiliations and the copyright notice.
% The command takes one argument, which is text to display at the start of the footnote.
% The \icmlEqualContribution command is standard text for equal contribution.
% Remove it (just {}) if you do not need this facility.

\printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
%\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

\begin{abstract}
    While fat-tailed densities commonly arise as posterior and marginal distributions in
    robust models and scale mixtures, they present challenges when Gaussian-based 
    variational inference fails to capture tail decay accurately. 
    We first improve previous theory on tails of Lipschitz flows 
    %\citep{jaini2020tails} 
    by quantifying how the tails affect the \emph{rate} of tail decay 
    and by expanding the theory to non-Lipschitz polynomial flows.
    We then develop an alternative theory for multivariate tail parameters which is 
    sensitive to tail-anisotropy. 
    In doing so, we unveil a fundamental problem which plagues many existing flow-based 
    methods: they can only model tail-isotropic distributions (i.e., distributions 
    having the same tail parameter in every direction).
    To mitigate this and enable modeling of tail-anisotropic targets, we propose 
    anisotropic tail-adaptive flows (ATAF).
    Experimental results on both synthetic and real-world targets confirm that ATAF 
    is competitive with prior work while also exhibiting appropriate tail-anisotropy.
\end{abstract}

@@

\bibliography{refs_ftvi}
\bibliographystyle{icml2022}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage
\appendix
\onecolumn

@@

\begin{figure*}[htbp]
    \centering
    \vspace{-.5in}
    \includegraphics[scale=0.8]{Figures/spiral.pdf}
    % \includegraphics[trim={0 0 9cm 0},clip]{Figures/spiral.pdf}\\
    % \includegraphics[trim={5cm 0 4.9cm 0},clip]{Figures/spiral.pdf}\\
    % \includegraphics[trim={9.3cm 0 0cm 0},clip]{Figures/spiral.pdf}\\
    \caption{Taking a tail-anisotropic distribution (left) and ``spinning'' it (middle) results in
        one-dimensional projections which oscillate between tail parameters (as seen in
        $\log p(\braket{X,e_0})$ in right panel) and result in an ill-defined
        direction-dependent tail parameter function $\alpha_X(\cdot)$ due to a
        divergent limit.
    }
    \label{fig:spiral}
\end{figure*}

>>>>

% \documentclass[nohyperref]{article}
\documentclass[thesis.tex]{subfiles}

% Recommended, but optional, packages for figures and better typesetting:
% \usepackage{microtype}
% \usepackage{graphicx}
% %\usepackage{subfigure}
% \usepackage{booktabs} % for professional tables

% % hyperref makes hyperlinks in the resulting PDF.
% % If your build breaks (sometimes temporarily if a hyperlink spans a page)
% % please comment out the following usepackage line and replace
% % \usepackage{icml2022} with \usepackage[nohyperref]{icml2022} above.
% \usepackage{hyperref}

% % Attempt to make hyperref and algorithmic work together better:
% \newcommand{\theHalgorithm}{\arabic{algorithm}}

% % Use the following line for the initial blind version submitted for review:
% % \usepackage{icml2022}

% % If accepted, instead use the following line for the camera-ready submission:
% \usepackage[accepted]{icml2022}

% % For theorems and such
% \usepackage{amsmath}
% \usepackage{amssymb}
% \usepackage{mathtools}
%     \usepackage{amsthm}

% % if you use cleveref..
% \usepackage[capitalize,noabbrev]{cleveref}

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% % THEOREMS
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \theoremstyle{plain}
% \newtheorem{theorem}{Theorem}[section]
% \newtheorem{proposition}[theorem]{Proposition}
% \newtheorem{lemma}[theorem]{Lemma}
% \newtheorem{corollary}[theorem]{Corollary}
% \theoremstyle{definition}
% \newtheorem{definition}[theorem]{Definition}
% \newtheorem{assumption}[theorem]{Assumption}
% \theoremstyle{remark}
% \newtheorem{remark}[theorem]{Remark}

% % Todonotes is useful during development; simply uncomment the next line
% %    and comment out the line below the next line to turn off comments
% %\usepackage[disable,textsize=tiny]{todonotes}
% \usepackage[textsize=tiny]{todonotes}

% % The \icmltitle you define below is probably too long as a header.
% % Therefore, a short form for the running title is supplied here:
% \icmltitlerunning{Fat--Tailed VI with Anisotropic Tail Adaptive Flows}

% % custom stuff
% \usepackage[bold]{hhtensor}
% \usepackage{mathtools}
% \usepackage{float}
% \usepackage{braket}
% \usepackage{dsfont}
% \usepackage{graphbox}
% \usepackage{pgf}
% \usepackage{comment}
% \usepackage{subcaption}
% \usepackage{enumitem}
% \input{../macros}

@@

% \twocolumn[
% \icmltitle{Fat--Tailed Variational Inference with Anisotropic Tail Adaptive Flows}

% % It is OKAY to include author information, even for blind
% % submissions: the style file will automatically remove it for you
% % unless you've provided the [accepted] option to the icml2022
% % package.

% % List of affiliations: The first argument should be a (short)
% % identifier you will use later to specify author affiliations
% % Academic affiliations should list Department, University, City, Region, Country
% % Industry affiliations should list Company, City, Region, Country

% % You can specify symbols, otherwise they are numbered in order.
% % Ideally, you should not use this facility. Affiliations will be numbered
% % in order of appearance and this is the preferred way.
% \icmlsetsymbol{equal}{*}

% \begin{icmlauthorlist}
% \icmlauthor{Feynman Liang}{yyy,fb}
% \icmlauthor{Liam Hodgkinson}{yyy,zzz}
% \icmlauthor{Michael W. Mahoney}{yyy,zzz}
% % \icmlauthor{Firstname4 Lastname4}{sch}
% % \icmlauthor{Firstname5 Lastname5}{yyy}
% % \icmlauthor{Firstname6 Lastname6}{sch,yyy,comp}
% % \icmlauthor{Firstname7 Lastname7}{comp}
% % %\icmlauthor{}{sch}
% % \icmlauthor{Firstname8 Lastname8}{sch}
% % \icmlauthor{Firstname8 Lastname8}{yyy,comp}
% % %\icmlauthor{}{sch}
% % %\icmlauthor{}{sch}
% \end{icmlauthorlist}

% \icmlaffiliation{fb}{Meta, Menlo Park, CA}
% \icmlaffiliation{yyy}{Department of Statistics, University of California, Berkeley, CA}
% \icmlaffiliation{zzz}{International Computer Science Institute, Berkeley, CA}
% % \icmlaffiliation{comp}{Company Name, Location, Country}
% % \icmlaffiliation{sch}{School of ZZZ, Institute of WWW, Location, Country}

% \icmlcorrespondingauthor{Feynman Liang}{feynman@berkeley.edu}
% %\icmlcorrespondingauthor{Firstname2 Lastname2}{first2.last2@www.uk}

% % You may provide any keywords that you
% % find helpful for describing your paper; these are used to populate
% % the "keywords" metadata in the PDF but will not be shown in the document
% \icmlkeywords{Machine Learning, ICML}

% \vskip 0.3in
% ]

% % this must go after the closing bracket ] following \twocolumn[ ...

% % This command actually creates the footnote in the first column
% % listing the affiliations and the copyright notice.
% % The command takes one argument, which is text to display at the start of the footnote.
% % The \icmlEqualContribution command is standard text for equal contribution.
% % Remove it (just {}) if you do not need this facility.

% \printAffiliationsAndNotice{}  % leave blank if no need to mention equal contribution
% %\printAffiliationsAndNotice{\icmlEqualContribution} % otherwise use the standard text.

% \begin{abstract}
    While fat-tailed densities commonly arise as posterior and marginal distributions in
    robust models and scale mixtures, they present challenges when Gaussian-based 
    variational inference fails to capture tail decay accurately. 
    We first improve previous theory on tails of Lipschitz flows 
    %\citep{jaini2020tails} 
    by quantifying how the tails affect the \emph{rate} of tail decay 
    and by expanding the theory to non-Lipschitz polynomial flows.
    We then develop an alternative theory for multivariate tail parameters which is 
    sensitive to tail-anisotropy. 
    In doing so, we unveil a fundamental problem which plagues many existing flow-based 
    methods: they can only model tail-isotropic distributions (i.e., distributions 
    having the same tail parameter in every direction).
    To mitigate this and enable modeling of tail-anisotropic targets, we propose 
    anisotropic tail-adaptive flows (ATAF).
    Experimental results on both synthetic and real-world targets confirm that ATAF 
    is competitive with prior work while also exhibiting appropriate tail-anisotropy.
% \end{abstract}

@@

% \bibliography{refs_ftvi}
% \bibliographystyle{icml2022}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \newpage
% \appendix
% \onecolumn

@@

\begin{figure*}[htbp]
    \centering
    \vspace{-.5in}
    \includegraphics[scale=0.8]{../Figures/spiral.pdf}
    % \includegraphics[trim={0 0 9cm 0},clip]{Figures/spiral.pdf}\\
    % \includegraphics[trim={5cm 0 4.9cm 0},clip]{Figures/spiral.pdf}\\
    % \includegraphics[trim={9.3cm 0 0cm 0},clip]{Figures/spiral.pdf}\\
    \caption{Taking a tail-anisotropic distribution (left) and ``spinning'' it (middle) results in
        one-dimensional projections which oscillate between tail parameters (as seen in
        $\log p(\braket{X,e_0})$ in right panel) and result in an ill-defined
        direction-dependent tail parameter function $\alpha_X(\cdot)$ due to a
        divergent limit.
    }
    \label{fig:spiral}
\end{figure*}
